\setcounter{page}{1} \pagenumbering{arabic}
\chapter{Introduction}

This monograph presents theoretical and empirical considerations in the field of entrepreneurship. The fundamental idea is that the act of entrepreneurship expresses a crucial development of the social division of labour; a phenomenon explicitly addressed during the 18th century. We specifically recall the work of Adam~\citet{Smith1776} who, in his discussion on the generation of economic wealth, places the operation and evolution of the division of labour at the forefront. Although Smith focusses on the economic activities of society from a dynamic point of view, his concept of the \emph{invisible hand}---and the theory that an economy is normally in some naturally \emph{stable state}---has dominated theoretical discourse. Over the last two centuries the dynamic concept of economic interaction that Smith sought to discuss in the \emph{Wealth of Nations} has been substituted for a more static analysis, and the importance that he placed on increasing returns to specialisation and the operation of the division of labour has been diluted. This is reflected in general equilibrium analysis, a cornerstone of traditional economic theory. The result of this analysis states that if there exists no government interference within the economic activities of individual agents then these activities bring about a stable situation that can be valued as a natural equilibrium in the economy.

The arrival of an economic system to a well-defined predetermined steady state, as expressed in general equilibrium modelling, contradicts the notion of active entrepreneurship. The act of entrepreneurship is a dynamic force that cannot emerge from a static economic steady state, as defined by general equilibrium theory; this is a perspective that is shared with more heterodox economic writers such as~\citet{Schumpeter1942}. As a consequence, the notion of the entrepreneur and the act of entrepreneurship lacks definition and application in traditional economic theory. The most compelling definitions of entrepreneurship emerge within more heterodox strands of economics and sociology, which subsequently lack an elaborate theoretical basis. The outcome is a shallow insight into the environmental context of entrepreneurship and the result that entrepreneurship has on the organisation and distribution of power within society.

The theory developed in this monograph expresses the importance of institutional structures and the organisation of trade to the process of entrepreneurship. We value the general equilibrium concept as an expression of a specific and efficient form of organisation of the trade processes in the economy. The assumptions enforced by the Walrasian model lead to the formation of a market whereby the axiom of perfectly competitive behaviour is expressed. The institutional structure is inflexible and the resulting organisation of society is perfectly formed, thus allowing no role for entrepreneurial actions. A more general way to consider the organisation of trade processes is through the perspective of a network, which emerges as a consequence of a pattern of exchanges between a population of agents under a given institutional system of governance and mechanism of trade. Under the Walrasian model of general equilibrium, a complete network emerges representing perfectly efficient market processes. In reality, networked markets are imperfectly connected; structural holes and opportunities for brokerage exist. These structural imperfections provide opportunities for entrepreneurial actors to exploit.

Throughout this monograph we propose other forms of organisation and trade processes in the economy such as markets with imperfect structure, costly interactions, and heterogeneous positions of economic agents that derive from specialisation and the social division of labour. We establish a fundamentally dynamic theory of the economy, thus allowing for the development of entrepreneurship in a natural way. This is the purpose of our initial theorising; it regards how we perceive the interaction space of the economy and allows us to appropriately define the entrepreneur and act of entrepreneurship. From this, we can analyse entrepreneurial action and power within the context of empirical examples. Overall, we feel that this provides a better description of how real-world economic processes take place. 

The monograph is divided into three parts, each of which discusses certain modelling aspects of the economy under different states, forms of trade organisation and entrepreneurial activities. In this introductory chapter, we propose the notion of a socially structured economy, our perception of conventional modelling techniques under the Arrow-Debreu model and a background into the theory of networked social structure and economic interactions.

\section{Socially structured economies}
\label{sec:SociallyStructuredEconomies}

Traditional economic theory is founded on the basis of \emph{methodological individualism}, \emph{methodological instrumentalism} and \emph{equilibration}. With methodological individualism, the individual decision-making agent has a central position in the theory. In most cases, an economy is described as a configuration of such individual economic agents, each of which attempts to optimise some well-defined pre-determined goal or objective. In this configuration it is assumed that the economic agent possesses the tools to optimise their utility; this is the crux of methodological instrumentalism \citep{Arnsperger2006}. This implies that economic decision processes in the economy are determined by the individual decision behaviour of these economic agents and by the configuration in which these economic agents are located. The outcome of this interaction is an equilibrium state represented as an organisational steady state. The economic system transitions from one steady state to another through some exogenous shock; suggesting that any dynamism is purely transitory and is fundamentally ill-defined---a \emph{manna from heaven}.

An additional assumption is that the individual economic agent can be described entirely with the use of individual attributes only. Examples of these individual attributes are an endowment bundle of commodities, individual production capacities and an individual preference relation on the set of well-defined, achievable commodity bundles for the particular economic agent. An illustration of this reliance on methodological individualism is with respect to the traditional Arrow-Debreu model of a perfectly competitive market system. This model contains a population of individual economic agents partitioned into a set of consumers and a set of producers. Consumers attempt to optimise a certain individual preference relation over the collection of achievable commodity bundles, while producers try to maximise their profit given their individual production capacities. These economic agents are placed within the context of a system of perfectly competitive markets, which puts certain constraints on the abilities of agents. The major constraint is given by the price levels determined by the market. Consumers limit themselves in terms of budget sets and, while producers adapt their production plan to maximise their profits. There is no consideration for other socially constructed characteristics that have an influence on the preferences and abilities of the individual decision-maker: the price level is the only `social' aspect that has a considerable influence on the population of individuals.

Within this context, the application of methodological individualism is very specific. Only individual decision-makers are participating in the processes that take place within the economy. Its use suggests that economic agents are essentially individuals who are interacting with each other according to the exogenous rules of the economic organisation and not with each other as a primitive social characteristic in their behaviour. This illustrates that the traditional method of microeconomic theorising implicitly involves a certain requirement on the `rules of the game'. 

In this monograph, we recognise that the rules of the game---which are reflected in the social and economic structure of the economy---are crucial in the design of a descriptive model of social and economic processes in that economy. Indeed, all economies are \emph{socially structured}; they are characterised by the socially constructed institutions and systems of governance that in turn guide the actions of a population of interconnected and interdependent agents. This social structure is flexible and, as such, is enforced and manipulated by economic agents themselves. This suggests that economic agents are not just individual decision-makers, but are decision-makers within a certain organisational environment consisting of a social and economic interaction architecture and behaviouristic conventions. An economic agent is still considered to be a decision-maker with some individual objective goal, but is also explicitly considered within some organisational environment determined by a well-defined governance system. The fact that an agent is considered as embedded within an organisational environment, leads to the acceptance of certain so-called \emph{social constraints} on the behaviour of that economic agent. Specifically, the agent's activity is contained within a matrix of interactions thus suggesting that the structure of interactions can also impose a constraint to their decision-making. We claim that constraints for some agents provide opportunities for others; distinctively, imperfections within society can provide opportunities for entrepreneurial activity.

\medskip\noindent \citet[p.~5]{Gilles1990} notes that the Arrow-Debreu model shows that objective market price expresses a specific form of aggregation---\emph{Walrasian aggregation}---which is a hypothesis of the organisation as represented by the market. Within the model, the price is representing the aggregation of demand and supply within the organisation of a perfectly competitive market. As a consequence, all agents communicate with each other through the market price only and not with each other directly. \citet{GillesDiamantaras2003} notes that this leads to a puzzle based on cyclical logic. But it more fundamentally suggests the inflexibility of perfect competition as an extremely efficient market structure. It is based on individualistic and symmetric economic agents that react to signals from market prices. The result is an equilibrium with no endogenous deviation and therefore has no capacity for wealth generation through entrepreneurial means. There are no deeper social features and attributes that influence the behaviour of the agents which shows that the functioning of the market is not influenced by other factors other than the market price.

The Arrow-Debreu model expresses a socially structured economy, albeit in a limited sense. The interactions of a finite set of individualistic decision-makers generate an idealistic market characterised by an objective price level for the exchange of a good. The socio-economic structure developed here is limited in the sense that it results in a static market mechanism shorn of all other social influences. In the real world, prices are not objective as in Arrow-Debreu style general equilibrium modelling. The notion of perfect competition is purely \emph{idealised}. Market imperfections, bounded human rationality, institutional structures and incomplete network architectures explain disparity and subjectivity in price levels. \citet{Blume2009} provide a model of exchange on a network based on the notion of Nash equilibrium. The model suggests that agents possessing dominant positions within a network of relations can influence price levels exposed to both producers and consumers. Models of this nature are discussed in Section~\ref{sec:socialeconomicnetworks} below. Institutions and socially constructed customs and cultures inform preferences and thus price levels within markets. We embrace these fundamentally social aspects of the economy within our modelling.

\subsection{Objectives}

The overall aim of the monograph is to provide a theory of entrepreneurship and the entrepreneurial function within socially structured economies. Our first objective in achieving this is to develop a perspective of embedded economic activity based on a population of specialised economic agents that form a functional social division of labour. The deepening of the social division of labour leads to a generation of wealth that benefits society as a whole. Specifically, we take a \emph{relational perspective} in which to view the foundations of social and economic phenomena. This foundation, established in Chapter~\ref{ch:relationalperspective}, provides a theory of social and economic interaction based on the use of socially embedded systems of governance which are successfully utilised given a foundation of trust. The relational perspective developed embraces Smith, yet deviates substantially from the traditional paradigm of the economics discipline founded on the Arrow-Debreu model.

The second objective is to explain and illustrate the role of the entrepreneur within the relational perspective and highlight the impact that entrepreneurship has on the evolution of the social division of labour, and therefore the development or, in some instances, the regression of the economy. Within our theory we note that the entrepreneur can have an impact on different dimensions of the economy: specifically impacting economic agents themselves, the economic interaction networks, and the institutions of the interaction environment.

The third objective is to provide a distinction between the act of entrepreneurship and the entrepreneurial function of the economy. The latter relates to the adaptive and creative act that many economic agents perform in producing outputs; this facilitates differentiation and great economic wealth, but not necessarily a new species of work or profession. The former relates to incredibly creative and disruptive acts that are reflective in the organisation of society and the interaction of new specialisations.

The final objective is to complement our theoretical discussion of the relational perspective and entrepreneurship with empirical analyses of entrepreneurial activities. This involves creating a set of tools that allows us to analyse data in a pragmatic way. From these tools, we can identify influential positions and activities within a networked economy. Our two main empirical applications are relate to the network of elite Florentine families and the directorate network of New York City during the beginning on the 20th century.

\medskip\noindent In developing a relational perspective of socio-economic activity we first provide a detailed insight into our central unit of analysis: the human agent. We investigate the micro-foundational elements of the human agent; elaborating on the agents' innate ability to form social relationships. We stress the co-evolution of the human brain with the growth of the increasingly complex social environment that it must navigate and organise. In doing so, we align our analysis with the hypothesis that the human brain has not developed in order to understand informational attributes of our environment but, converse to the conventional wisdom, our brains still remain imperfect in understanding environmental information. The economic decisions we make in complex situations are only made on the basis of bounded rationality, with informational constraints, and in situations of uncertainty.

Our characterisation of the economic agent relaxes two important assumptions of traditional economic modelling. The first is that there exists a dichotomy between consumers and producers: the population of economic agents is partitioned into two non-intersecting sets, each of which interact between, but not within, one another. The second refers to the structure of production possibilities, which do not lend themselves to Smith's notion of increasing returns to specialisation. As a consequence, specialised producers do not emerge naturally within the Walrasian framework. In modelling economic agents, we allow for increasing returns to specialisation through learning, which stands in contrast to the Arrow-Debreu model and other general equilibrium models developed from its foundation.

From this perspective of the human actor, and the society in which she is embedded, we elaborate on a theory that takes a topological analysis of social and economic interactions through the evolution of institutions that facilitate exchange and the deepening of the division of labour. We suggest that the division of labour naturally leads to the formation of relational networks, and from this we investigate the evolution of these relational networks---and thus the evolution of the economy---through the perspective of the entrepreneurial function, which primarily acts to extend the division of labour. As such we contend that the evolution of the socio-economic networks, and the systems of governance that hold them together, can only be explained through the use of the entrepreneurial function and the subsequent extension, or contraction, of the social division of labour.

\subsection{Motivation}

The development of a novel model for investigating socio-economic phenomena has come from two overlapping concerns. The first is prompted by the apparent deficiency in mainstream theory to convincingly address the emergence of new organisational structures and technologies and to fully incorporate the entrepreneurial function within the economy. The entrepreneur is left undeveloped in many formal economic theories. The main problem of these mainstream theories lies in appropriately modelling the entrepreneurial function derives from the use of Walrasian axioms based on the notion of general equilibrium as expressed above. Economic markets tend to converge to a predetermined equilibrium state within which unique market-clearing prices for homogeneous outputs emerge. At such a point it is difficult to endogenously integrate the development of an innovative technology into the model without having to assume its eventual existence which exogenously disrupts the growth of the economy. Well-established endogenous growth models have been created to formally explain how the generation of ideas from a research sector can lead to economic growth~\citep{Romer1990}, however, these models do not truly integrate differentiated and innovative commodities, nor do they convincingly integrate the entrepreneur and innovative entrepreneurial actions. In order to integrate the entrepreneurial function, it may be beneficial to maintain that the notion of static equilibrium as a purely intellectual concept. Much like the Austrian perspective on the matter, opportunities arise when the economy is imperfect and out of equilibrium; and, much like the Schumpeterian perspective, the entrepreneur's actions actively \emph{disequilibriate} the economy thus facilitating the emergence of increasingly more opportunities and innovation to take place meaning that equilibrium will never emerge. Thus, we should advocate that the economy has the potential to be in a continual state of dynamism.

Second, and perhaps more importantly, there seems to be an underlying disgruntlement within the economics discipline which has especially come to light in both academic and popular literature since the global financial crisis of 2008. Some economists believe that the fundamental problem derives from the fact that the economics discipline has become progressively divorced from economic reality.~\citet{Hodgson2009},~\citet{Smith2010} and~\citet{Keen2011} have appealed for a complete reform of the economics discipline as opposed to continually building increasingly esoteric and irrelevant mathematical models on top of a broken paradigm. These critics agree that the focus of the discipline has to change from its overwhelming focus on elegant General Equilibrium models that provide highly trivial results to a more pragmatic and empirically founded approach based on inductive reasoning. Where the discipline should concentrate its efforts remains debatable and highly sensitive to the subjective opinion of the academic economist. We agree with Hodgson that, in order to be pragmatic, we must orient ourselves to understanding real-world institutional environments and actors. However, it is important to initially be more fundamental than that; we must first understand the mechanisms concerning the embeddedness of economic activity in social constructs, the implications of the resulting division of labour and trade, and the development of innovation which is at the heart of the entrepreneurial function.

This disgruntlement concerning the state of the economics discipline is far from new. Nobel prize winners, such as Wassily Leontief, Milton Friedman and John Hicks have expressed their concern. Ronald~\citet{Coase1997} was explicit in claiming that, ``[e]xisting economics is a theoretical system which floats in the air and which bears little relation to what happens in the real world... I would say it has no subject matter. That's the problem.''~\citet{Knight1921} opened his influential book, \textit{Risk, Uncertainty, and Profit}, by highlighting the problems of theoretical economics perceiving itself as the only social science aspiring to become an exact science. He suggests that the discipline has provided itself with a seemingly impossible task of identifying and anatomising extremely complicated and intertwined social and economic phenomena, which perhaps cannot be measured. Even at that time, Knight highlighted the problems of divorcing a single individual from the entirety of the structure the interact in, a problem that is still expressed.~\citet{Krugman2009},~\citet{Stiglitz2010} and~\citet{Varoufakis2011} have expressed their concerns regarding the disciplines divorce from reality and its politically-driven ideology. The discipline may be better categorised as a subjective art---or even in some cases as a religion---as opposed to a true and exact science~\citep{Backhouse2010}.

\medskip \noindent This monograph provides the foundations of a coherent and productive paradigm; a holistic perspective in which to frame social and economic phenomena. The relational perspective that is advocated revolves around a number of interdependent microeconomic concepts: institutions, increasing returns to specialisation, the division of labour, and entrepreneurship. Using the foundational concepts of the relational perspective we explain the entrepreneur and the entrepreneurial function within the economy, therefore investigating innovation and the emergence of new institutional and organisational structures.

This discussion of the entrepreneur is not one that considers a single economic agent exogenously endowed with entrepreneurial talent or ability within a social and institutional vacuum as is considered in much of psychology. The perception of the entrepreneur in this monograph is a story of economic institutions, social structure, and the positions and actions of individuals within society. Through the analysis of this revolutionary character we can also tell a story of dynamic change within both the technological and institutional spheres of the economy, as well as a story on the current state of the institutions that facilitate and guide the entrepreneur and the division of labour, and how they may be improved to allow for progressive social and economic development.

\section{The relational perspective}

Economic agents are innately social. As such, all economic activity is socially embedded. We wish to introduce these social characteristics into the description of the economic agent and the environment they operate in. This is done in a direct manner in that we describe social characteristics as potential pairwise relationships between economic agents that specifically refer to the social and economic activities of these agents. The most natural of these relationships is a potential economic exchange relationship; two agents, who are related in a fundamentally social way, may activate their trade relation if it is mutually beneficial to do so.

The introduction of these social characteristics in the form of pairwise relationships leads to the formation of a global social structure on the population of economic agents. These social characteristics are not individualised, but should instead be seen as \emph{relational} as they must involve more than one economic agent in order to exist. In following~\citep{Gilles1990} we note that the relations between economic agents can be treated separately from the individual characteristics of those agents. The application of this modelling principle allows us to talk about the network of relations independently from the characteristics of the agent. As such, we apply graph theory to the analysis of the relational perspective. An introduction to the theoretical and empirical debates on social and economic networks are given in Section~\ref{sec:socialeconomicnetworks} below.

The evolving relational structure places social constraints on each individual economic agent. These social constraints come from two phenomena. First, from the relational structure of social and economic networks that the agents develop through social relationships and exchange. Second, from the socially constructed institutions and systems of governance that guide agents through the formation of relational networks. The existence of systems of governance brings with it a rule set that constrains the potential actions of individual agents and allocates resources between the population of economic agents. Both social constraints are investigated throughout the monograph.

The recognition and analysis of social structure and social constraints is not new. The investigation to how participants act within a social economy was one of the eventual objectives of~\citet{vNM}. This general research into the interaction of social structure and economic outcomes was extended by~\citet{KalaiMiddlemen1978} who remarked that social defects within an economic environment can lead to an uneven distribution of power within society. Specifically, there will exist an unequal distribution of power throughout society. Indeed, intermediating agents can potentially leverage their position within a matrix of relationships, exercising their power on others agents that they are connected to. An interesting point is raised---and one that is noted in other works since \citet{Granovetter2005}---is that the social abilities of a set of agents can have a significant impact on their economic outcomes~\footnote{Although the economic benefits of dominant middleman activities have been substantially highlighted, \citet{KalaiMiddlemen1978} focus also on the costly disadvantages of sustaining a middleman position. Specifically, through a comparison of the Core, the authors note that the position of a middleman can be burdensome.}.

\subsection{Middlemen and entrepreneurs}

The notion of a middleman is introduced in conjunction with that of a network. We define a middleman as a graph-theoretical concept and embed the notion of power---in terms of social and economic connectivity---within the context of \emph{centrality}. The philosophy behind this is to provide an objective measurement to economic power within a trading platform given some social structure. From this, we can derive the distribution of utility through the network of socio-economic relationships. Indeed, the naturally forming network configuration represents the aggregate utility and the distribution of utility throughout society. These networked economies and the presence of middlemen can only be explained fully within a relational perspective of socio-economic activity.

This monograph combines the notion of a middleman with that of the entrepreneur. Specifically, we claim that the middleman position of an economic agent represents the existence of an entrepreneurial activity that has lead to a new specialisation, and thus an extension of the division of labour. The formation of a middleman position---whether it be through individual efforts or group activity---is illustrative of entrepreneurship and reflective of a changing interaction infrastructure.

We use numerous methodologies to analyse entrepreneurship and power within the relational perspective including graph theory, \emph{non-cooperative} game theory and statistical analysis. Specifically, we use graph theory in identifying middleman and providing an objective measurement of power; non-cooperative game theoretic concepts to explain the formation of middleman positions within networked systems, and statistical analysis to measure the significance of an agents middleman position and centrality to their economic performance. This investigation into the notion of the middleman, and its relationship with the entrepreneur is made after first establishing the relational perspective.

\section{An introduction to social and economic networks}
\label{sec:socialeconomicnetworks}

Much of the formal analysis of the relational perspective is based on network theory. This section provides initial definitions and insights into network theory and the literature around social and economic networks. Although some of this provides only a background to network analysis, many of the concepts discussed below are used more formally in subsequent chapters.

\subsection{Basic concepts}

A network is a graph with non-trivial topological features such that there exists a set of \emph{nodes}---which represent a corresponding set of social actors or economic agents---that are connected to one another in some fashion by a set of \emph{links}---that represent relationships---whereby each link can connect nodes either directly or indirectly. Nodes can be connected to each other by a \emph{path} if from one node to another there exists a set of links that can be traversed such that one node can manoeuvre to the other. The shortest path between any pair of nodes is termed as a \emph{geodesic path}, and the number of links that need to be traversed along this geodesic path is termed as the \emph{geodesic distance}. A \emph{strongly connected network} is one in which all nodes can connect to each other either directly or indirectly by a path.

Generally, it's found that many networked structures that emerge have topological properties that are neither purely random nor regular. Research into neural networks has highlighted a commonality regarding the complexity and networked structure of seemingly unrelated social, economic, and natural systems \citep{SpornsTononi2005, Sporns2010}. When modelling the structure of social networks \citet[p.~440]{WattsStrogatz1998} noted that, ``these systems can be highly clustered, like regular lattices, yet have small characteristic path lengths, like random graphs''. Many complex networks exhibit this mixture between random and non-random wiring whereby a few small dense clusters of actors are connected together by a few weak relationships.

A scale-free network, formally developed by \citet{BarabasiAlbert1999}, is one in which the number of links that an actor has in the network is unequally distributed such that there exists a minority of actors with many links and a majority of actors with relatively few links. Actors with relatively many links are termed as \emph{hubs}; the presence of hubs substantially reduces the diameter of the network.

\subsection{Characteristics of social networks}

\citet{JacksonRogers2007} summarise some key empirical regularities shared by networks created by social actors. First, they note that the average geodesic distance between any pair of nodes in the network has a tendency to be small, which corresponds well with the above finding of densification, and the maximum distance between any pair of nodes in a social network is small. This observation that there may exist a short distance between any two individuals, or nodes, was initially noted by Hungarian playwright Frigyes \citet{Karinthy1929} who conducted a thought-experiment regarding how close two random people were to each other by their local connections. Also, sociologist Georg Simmel described modern society as consisting of loosely connected social circles of relationships \citep{Simmel1950}.

This thought-experiment was operationalised and later carried out in a more scientific way in the form of the \emph{small-world experiment} by Stanley \citet{Milgram1967}. This experiment was based on the methodology whereby a letter sent from a city in the midwest of the USA to Boston, USA, by means of transfer between personal acquaintances. The small-world experiment was designed to measure these path lengths by developing a procedure to count the number of links between any two acquaintances in the corresponding chain. The experiment resulted in the fact that the average observed path length was between 5.5 and 6, resulting into the well-known phrase ``Six Degrees of Separation''. This experiment has been replicated across different countries and with different techniques and the general finding by Milgram has been maintained; thus there tends to exist a low distance between any pair of nodes relative to the size of the population as a whole.

Second, Jackson and Rogers note that the clustering coefficient of the network---which measures the tendency of connected nodes to have common neighbours---are larger in social networks compared to networks where the links are randomly generated between nodes. This finding suggests that communities are formed in the network whereby nodes who share a given personal, geographical, or institutional characteristic tends to form a link with each other \citep{Watts2002}. The aggregation of these links across a number of nodes forms a clique, such that all nodes within a clique are connected to each other. \citet{KossinetsWatts2006} note that \emph{focal closure}---whereby people who are members of the same affiliation have a tendency to form a link with each other---as a reason for why we tend to see this clustered pattern of communities on a network. This finding is associated with the notion of social foci, developed by Scott \citet{Feld1981, Feld1982}, as organisations that induce high connectivity between members.

Third, the distribution of the number of links a node has, i.e., a nodes' \emph{degree}, tends to exhibit fat tails so that there are more nodes with relatively high and low degrees and fewer nodes with medium degrees than one would find if the links were formed in a random way. The degree distribution can be claimed to be approximately \emph{scale-free} such that the distribution follows a power-law. People who are more popular in terms of the number of connections they have are seen to be more attractive to form a relationship with, thus inducing this rich-gets-richer phenomenon initially found by \citet{Simon1955}.

Fourth, there tends to exist some positive \emph{assortativity} such that the degrees of connected nodes tends to be positively correlated in that nodes with a large degree tend to be connected to other nodes with a large degree, and nodes with a low degree tend to be connected to each other \citep{Newman2003mixing}. This assortativity between nodes extends beyond topological properties of networks such that nodes with a certain personal characteristics form connections between each other; this is a notion, distinct from assortativity, is known as \emph{homophily}. These notions of assortativity and homophily are initial findings of the context of social networks in sociology \citep{McPherson2001} and remain interesting topics in medicine and finance \citep{Haldane2009, HaldaneMay2011}.

Finally, Jackson and Rogers note that the clustering among neighbours of a given node tends to be inversely related to the nodes degree. This final point is highly intuitive---especially given the notion of positive assortativity between nodes in a network---as the probability of all of a well-connected nodes neighbours to be connected to each other would be low.

\subsubsection{Models of social networks}

Since noting these characteristics a variety of random graph models have been proposed to explain some of the characteristics. The three most common models are that of a \emph{random network}, a \emph{small-world network}, and a \emph{scale-free network}. The notable structures of small-world and scale-free networks can be seen in Figure~\ref{networks} (a) and (b) respectively. These models of network formation can be either random and algorithmic, or can be strategic with the use of game theory whereby individual agents have some incentive to either form or sever links.

\begin{figure}[t]
\begin{center}
\begin{tikzpicture}[scale=0.5]
\draw[thick] (0,5) -- (5,0);
\draw[thick] (10,5) -- (5,0);
\draw[thick] (5,10) -- (10,5);
\draw[thick] (5,10) -- (0,5);
\draw[thick] (1.5,1.5) -- (1.5,8.5);
\draw[thick, dashed] (1.5,8.5) -- (8.5,8.5);
\draw[thick] (1.5,8.5) -- (8.5,1.5);
\draw[thick] (8.5,8.5) -- (8.5,1.5);
\draw[thick] (8.5,1.5) -- (1.5,1.5);
\draw[thick] (0,5) -- (1.5,1.5);
\draw[thick] (1.5,1.5) -- (5,0);
\draw[thick] (5,0) -- (8.5,1.5);
\draw[thick] (8.5,1.5) -- (10,5);
\draw[thick] (10,5) -- (8.5,8.5);
\draw[thick] (8.5,8.5) -- (5,10);
\draw[thick] (5,10) -- (1.5,8.5);
\draw[thick] (1.5,8.5) -- (0,5);

\draw[thick] (17.5,0) -- (22.5,0);
\draw[thick] (17.5,0) -- (15,5);
\draw[thick] (17.5,0) -- (20,2.5);
\draw[thick] (22.5,0) -- (20,2.5);
\draw[thick] (22.5,0) -- (25,5);
\draw[thick] (25,5) -- (20,2.5);
\draw[thick] (25,5) -- (20,10);
\draw[thick] (25,5) -- (21,7);
\draw[thick] (20,2.5) -- (19,7);
\draw[thick] (20,2.5) -- (15,5);
\draw[thick] (15,5) -- (20,10);
\draw[thick] (20,2.5) -- (21,7);
\draw[thick] (19,7) -- (15,5);

\draw (0,5) node[minimum size=2em,circle,fill=red!20] {$1$};
\draw (5,0) node[minimum size=2em,circle,fill=red!20] {$2$};
\draw (5,10) node[minimum size=2em,circle,fill=red!20] {$3$};
\draw (10,5) node[minimum size=2em,circle,fill=red!20] {$4$};
\draw (1.5,1.5) node[minimum size=2em,circle,fill=red!20] {$5$};
\draw (1.5,8.5) node[minimum size=2em,circle,fill=red!20] {$6$};
\draw (8.5,1.5) node[minimum size=2em,circle,fill=red!20] {$7$};
\draw (8.5,8.5) node[minimum size=2em,circle,fill=red!20] {$8$};
\draw (17.5,0) node[minimum size=2em,circle,fill=blue!20] {$1$};
\draw (22.5,0) node[minimum size=2em,circle,fill=blue!20] {$2$};
\draw (15,5) node[minimum size=2em,circle,fill=blue!20] {$3$};
\draw (25,5) node[minimum size=2em,circle,fill=blue!20] {$4$};
\draw (20,2.5) node[minimum size=2em,circle,fill=blue!20] {$5$};
\draw (20,10) node[minimum size=2em,circle,fill=blue!20] {$6$};
\draw (19,7) node[minimum size=2em,circle,fill=blue!20] {$7$};
\draw (21,7) node[minimum size=2em,circle,fill=blue!20] {$8$};

\draw (5,-2) node {(a)};
\draw (20,-2) node {(b)};
\end{tikzpicture}
\end{center}
\caption[Small-world and scale-free networks]{Network (a) shows a circular small-world model and network (b) shows a scale-free model.}
\label{networks}
\end{figure}

\paragraph{Random graph models.}

The first set of network formation models to be developed are based on the random generation of links between nodes. A random graph is a general term that refers to probability distributions over graphs. Random graphs may be described by a probability distribution, or by a random process which generates them. They are obtained by having an initial set of unconnected nodes and then randomly connecting nodes together by a set of links. The initial research into random graph models was conducted by \citet{ErdosRenyi1959}. The seminal model developed serves as the basis for all subsequent random graph models \citep{Gilbert1959, Bollobas2001}.

A number of properties are derived from the random graph models. An important property is that of a \emph{phase transition}. A phase transition refers to a punctuated change in the state, structure and dynamics of a complex system, that arises from a slight change in a variable that characterises the system as a whole. In the context of an Erd\"{o}s-R\'{e}nyi graph, if the probability of nodes being connected falls below a certain threshold the network will most certainly be \emph{disconnected}, such that there will exist at least one node that is not connected---at least indirectly---to other nodes and thus there cannot exist a strongly connected network. Above the threshold then the network will be strongly connected. The phase transition from a connected network to a disconnected network thus occurs when the probability of nodes being connected falls below a well-defined level.

Although theoretically compelling, many social and economic networks are not formed `randomly'. For example, empirical social networks are much more clustered than those generated by random. Real-world networks are neither completely ordered nor completely random, but rather exhibit important properties of both. Models of small-world networks add `order' to the network by assuming a regular lattice as the starting point.

\paragraph{Small-world networks.}

A small-world network is one whereby nodes are not necessarily neighbours of one another, but most nodes can be reached from every other by a small number of steps. In this form of network, \citet{WattsStrogatz1998}, \citet{Watts1999}, and \citet{NewmanStrogatzWatts2001} generate networks by starting with a symmetric, regular network and randomly rewiring some links between nodes. They found that the random rewiring process leads to an exponential decrease in the average distance between any two nodes whilst maintaining a network that is much more clustered than just randomly wiring the network. Figure~\ref{networks} (a) effectively shows the Watts-Strogatz model of small-world networks. In this example, a regular lattice is generated on a set of eight nodes whereby each node has four links each. Node 6 is rewired to another randomly, which reduces the average path length of the network.

\citet{Kleinberg2000a, Kleinberg2000b} also provides an algorithmic approach to modelling the small-world phenomenon in which he develops a decentralised algorithm to derive the efficient amount of randomness required to generate a small-world network. In the economics literature, \citet{JacksonRogers2005} provide a game theoretic approach to the formation of pairwise stable small-world networks where agents wish to maximise the amount of information they have access to either directly or indirectly given that the information flowing through the network depreciates the longer the geodesic path from one node to another.

\paragraph{Scale-free networks.}

A scale-free network is one in which the number of links that an actor has in the network is unequally distributed such that there exists a minority of actors with many links and a majority of actors with relatively few links. Actors with relatively many links are hubs whose presence reduces the diameter of the network. Research by \citet{Price1976}, \citet{BarabasiAlbert1999}, and \citet{CooperFrieze2003} has shown that networks with power law degree distributions result if nodes form links through the process of \emph{preferential attachment} (i.e., new nodes link to existing nodes with probabilities proportional to the existing nodes' degrees). Power-law distributions have also been shown to result if new nodes copy the links of a randomly identified node \citep{Kleinberg1999, Kumar2000}, or if networks are designed to optimise tolerance \citep{Fabrikant2003}. A variation on preferential attachment where only some nodes are active at any time \citep{Klemm2002a} has been shown to also exhibit small world properties. Some network models that grow over time have been shown to exhibit assortativity and scale-freeness, e.g., \citet{Callaway2001} and \citet{Krapivsky2002}.

Research has shown that many social and economic networks share a scale-free property \citep{Faloutsos1999, DorogovtsevMendes2003, Barabasi2009, Barabasi2012}. Also, biological and mechanical networks also share this structure \citep{Barabasi2011}. Scale-free networks have a number of desirable qualities including robustness to random attacks and small diameters \citep{AlbertJeongBarabasi1999, AlbertJeongBarabasi2000}. As a consequence, scale-free networks are prone to collapse due to targeted attacks and contagious processes.

While the aforementioned models result in some of the empirical regularities of large social networks, none of them are consistent with all of the characteristics above. The closest network formation process to cover all of these properties is indeed from \citet{JacksonRogers2007} whereby agents enter the network and initially form a network randomly, then create other links to nodes based on the initial connections' links. This formation process---a hybrid between random and strategic network formation---provides a structure that is most resembling real social networks. However, it is still questionable as to whether the methods of generating these `social' networks are actually resembling the processes underlying most of the large networks that we actually observe.

For the moment we are only interested in the structure and the methods of forming networks. In Chapter~\ref{ch:relationaltheory} we consider the formation of exchange networks under an algorithmic process. We find that depending on the method of exchange between nodes the overall structure of the network and the average utility of the population changes. The generation of these networks serves as a basis for the algorithmic generation of economic relationships.

\subsection{Economic networks}

There are many studies of markets as networks \citep{RauchCasella2001}. \citet{Uzzi1996} provides one of the most influential investigations regarding the importance of social relationships in the New York City clothing industry. He focussed on the type of relationship between the clothing firms. Specifically, he categorised relationship as ``market'' or ``arm's length'', which included one-time social and economic interactions; and ``special'' or ``close'' interactions, which included many relationships with repeated interactions. Uzzi found that three main aspects were important for these relationships and interactions, these are joint problem-solving; trust; and information transfer. From his analysis, he found that firms with more clustered and intertwined social and economic relationships have a greater probability of survival. To Uzzi, social and economic embeddedness helps firms to survive and thrive. Also, the study found the importance and differences of market economic and close social relationships.

\citet{Weisbuch2000} and \citet{KirmanVriend2001} consider the Marseille Fish Market where market prices are not posted publicly. This allows prices to be maximally flexible: Each producer decides about her own prices; different consumers might be set different prices, and prices can vary over time. Prices are communicated privately and there is no bargaining in this market. A number of outcomes emerge. First, there is widespread loyalty to a single producer, while the minority of consumers purchase from multiple producers. The market seems promotes trust and reciprocity between consumers and producers. Specifically, consumers follow an adaptive updating rule, with a higher tendency to visit producers who have met their demands in the past. \citet{Vignes1993} and \citet{VignesEtienne2011} assess the network structure and the distribution of price levels throughout the market. The trade pattern that emerges has an interesting feature in that there is persistent price dispersion that fits with the networked structure of the market. The network tended to follow a scale-free structure.

\citet{Granovetter1973} famously considered the labour market a social network of references. As part of his PhD thesis, he investigated how people within a certain professional socio-economic class typically found information regarding new job opportunities. He identified that people tended to obtain the most relevant information about job availability from contacts in their social neighbourhood. Less intuitively, he found that the best information often came from acquaintances as opposed to close friends or family. Granovetter claimed that every person who forms social relationships holds a mixture of both \emph{strong} and \emph{weak} ties; where strong ties pertained to close friends and family members, and weak ties pertained to acquaintances. According to Granovetter, the labour market operates as a social network where information regarding job opportunities are disseminated through the network. Weak ties, as opposed to strong ties, held less redundant and therefore more diverse information which an individual could exploit when attaining a job. Much in the same vein as Granovetter, \citet[p.~333]{Jackson2008} suggests that social networks of a firms' existing employees can also provide a good resource for searching for suitable employees, `` [a] firm may simply want to hire people similar to the employees it already has. Given the homophily in many social networks, a firm can take advantage of its existing workforce to find other people with similar characteristics.'' The outcome of the assessment is that the labour market that exists is more intricate than one characterised by traditional microeconomic analysis of labour economics.

The theory of weak and strong ties has been the cornerstone of much of economic sociology since Granovetter's initial findings. The weakest, and therefore most valuable in terms of information, relations are \emph{bridge} relations. These were discussed in depth by \citet{Burt1992}, who suggested that individuals could exploit these bridge relationships through the attainment of new and diverse sets of information \citep{Burt2004}. As such, there exist positions of power within social networks and networked markets where some positions are more influential than other positions.

Exchange theory is concerned with how the structure of relationships among agents affects exchange between them \citep{CookWhitmeyer1992}. These exchanges could include economic interactions, trading of favours, communication of information, or other social interactions. Typically this research is based on dyadic interactions, but since the work of Robert \citet{Emerson1962, Emerson1972a, Emerson1972b, Emerson1976} networks have played an increasingly important role in exchange theory. In his work, Emerson considered network interactions to explain power and dependencies when considering an exchange. This theory of networks played a role in Emerson's work because all economic exchange depended on the availability of outside options; therefore assuming a relatively competitive market in which economic relations could be made and severed without any friction. Under such a perspective, an economic interaction cannot be viewed without the context of the complete network of all potential interactions and relationships. Again, much like Burt, network positions matter and asymmetries of power tend to be expected.

\subsubsection{Models of economic networks}

Emerson's work combined with the mathematical research of matching in bipartite graphs \citep{Hall1935} provided the basis for more research in bilateral trading models \citep{Corominas-Bosch1999, Corominas-Bosch2004}. Much like the implicit assumption of general equilibrium models, these models assume a dichotomy between the demand and supply within an economy. Therefore, all networks that emerge are simply between a set of nodes who produce the supply of the economy and a set of nodes who demand the produced goods. \citet{Blume2009} provide a convincing model of economic exchange on networks. They also dichotomise the forces of demand and supply by introducing a set of `consumers' and a set of `producers' whereby a consumer can only engage in an economic exchange with a producer. Each economic exchange is represented as a link. As a consequence, a bipartite network is formed whereby economic exchange occurs across the links between consumers and producers. The structure of their model follows that of \citet{KrantonMinehart2001} whereby, again, the forces of demand and supply are partitioned therefore forming two disjoint sets of consumers and producers. This bipartite set-up is also taken by \citet[Chapter~10]{Jackson2008}. Easley and Kleinberg find that this networked structure provides some nice properties, being able to show that society's welfare is maximised when prices develop to guide consumers to producers. Moreover, they find that, along with \citet{Kakade2004a, Kakade2004b}, a network provides a good basis for price dispersion as was seen in the Marseille fish market case.

\begin{figure}[t]
\begin{center}
\begin{tikzpicture}[scale=0.5]
\draw[thick] (5,0) -- (15,5);
\draw[thick] (10,0) -- (5,5);
\draw[thick] (20,0) -- (10,5);
\draw[thick] (15,0) -- (25,5);
\draw[thick] (25,0) -- (20,5);

\draw (5,0) node[minimum size=2em,circle,fill=green!20] {$6$};
\draw (5,5) node[minimum size=2em,circle,fill=orange!40] {$1$};

\draw (10,0) node[minimum size=2em,circle,fill=green!20] {$7$};
\draw (10,5) node[minimum size=2em,circle,fill=orange!40] {$2$};

\draw (15,0) node[minimum size=2em,circle,fill=green!20] {$8$};
\draw (15,5) node[minimum size=2em,circle,fill=orange!40] {$3$};

\draw (20,0) node[minimum size=2em,circle,fill=green!20] {$9$};
\draw (20,5) node[minimum size=2em,circle,fill=orange!40] {$4$};

\draw (25,0) node[minimum size=2em,circle,fill=green!20] {$10$};
\draw (25,5) node[minimum size=2em,circle,fill=orange!40] {$5$};

\draw (0,0) node {Producers};
\draw (0,5) node {Consumers};
\end{tikzpicture}
\end{center}
\caption{A matching market of consumers and producers}
\label{fig:bipartiteexchange}
\end{figure}

This bipartite representation of a market, known as a ``matching market'', can be seen in Figure~\ref{fig:bipartiteexchange}. This notion of a matching market is further extended by \citet{EasleyKleinberg2010} to include `traders' as intermediaries between consumers and producers. From this they informally define a notion of competition in a networked economy, suggesting that traders have an opportunity to monopolise flows of economic exchange between producers and consumers. The notion regarding the monopolisation of trade is extended and considered in detail in Chapters~\ref{ch:criticalnodes} and~\ref{ch:blocks} where we discuss measurements of brokerage in networked economies.

The reason why networks follow a bipartite structure is due to the way in which economic agents are characterised within the traditional perspective, and specifically the strict dichotomy between demand and supply~\footnote{When generating economic networks in Chapter~\ref{ch:relationaltheory} we do not assume the existence of a strict dichotomy between consumption and production and, as a direct consequence, a bipartite network does not typically emerge. Instead, when generating the exchange networks we arrive at properties that are more closely related to the typical structure of social networks as noted by the characteristics above.}. Within this monograph, we note that economic agents are boundedly rational and exist within a network of social and economic relations that are influenced by systems of governance. However, the economic agent is further defined such that each individual is endowed with the powers of both production and consumption; that the forces of demand and supply are contained in all individual economic agents.

\section{Monograph structure}

This monograph is partitioned into three interlinked parts, each of which addresses the ultimate goal of developing a theory of the entrepreneur and entrepreneurial activity within a relational perspective of socio-economic activity. The first two parts are primarily focussed on theory, occasionally using empirical examples to illustrate and test models and measures. The final part is more focussed on explaining entrepreneurial activities within real-world case studies. Part~\ref{part:generalTheoryEntrepreneurship} consists of three chapters that provide a theoretical foundation to the relational perspective and an introduction to the entrepreneur. Much effort is spent developing the relational perspective framework, which is done from an evolutionary basis of the social actor that in turn represents an economic agent. Within this framework we are able to provide a holistic definition of the entrepreneur, linking the notion appropriately to the social division of labour and the networked structure that economic interaction produces. Of specific interest are the unique positions within the network that the entrepreneurial agent operates, and can be a source of their economic utility.

Part~\ref{part:entrepreneurshipPlatonianEconomy} consists of two chapters that elaborate on the interaction between entrepreneurship and the network interaction infrastructure. We investigate the establishment of uncontested middleman positions and the formation of blocks in relational economies. In elaborating on these concepts we develop formal models and measures to use on network-based data sets. These models and measures are then tested against the empirical data.

Part~\ref{part:entrepreneurshipPlatformEconomy}, the final Part of the monograph, also consists of two chapters that introduce an elaborate example of power within a more developed economy. This economic space is representative of a more layered economy within which there exist firms and directors that operate in multiple complementary industries. We analyse the directorial network and interlock formation in New York City from 1902--1912 as an entrepreneurial process. Indeed, these chapters investigate the establishment of power in New York City through network entrepreneurship and provide measures and models to investigate the centralisation of economic influence and use this to challenge claims regarding the history of the so-called \emph{Empire State} and the role of J.P. Morgan.

